ML Estimation of Matrix--$t$ Distributions with Expectation Maximization

### 2017-01-20 paper.tex
Adjusted the E-step expressions.

We can split the integral in Eq.~\eqref{eq:expdef} by the terms of $\log f(T_{j},S;\theta)$. Eliminating a global factor $\frac{1}{2}$ and terms which do not depend on the parameters in $\theta$ we have
\begin{equation}
\begin{split}\log f(T_{j},S;\theta)\sim -\nu\log|S|-\textrm{tr}\left[S^{-1}(\Sigma+(T_j-M)\Omega^{-1}(T_j-M)^{\textrm{t}})\right]\\
-\nu p\log2-p\log |\Omega|+(\nu+p-1)\log |\Sigma|-2\log\Gamma_{p}\left(\frac{\nu+p-1}{2}\right).
\end{split}\label{eq:logf}
\end{equation}
From Eq.~\eqref{eq:logf}, we can see the only integrals which need to be computed are
\begin{equation}
\int_{\mathbf{S}_+^p}f(T_j,S;\theta_k)\log |S|\,dS\label{eq:firstlogpart}
\end{equation}
and
\begin{equation}
\int_{\mathbf{S}_+^p}f(T_j,S;\theta_k) S^{-1}\,dS.\label{eq:2ndlogpart}
\end{equation}
We note that taking the derivative with respect to $\alpha$ in Eq.~\eqref{eq:intmultgammadef} we have
\begin{equation}
\int_{\mathbf{S}_+^p}e^{-\textrm{tr}\left[SA\right]}|S|^{\alpha}\log |S|\,dS=\Gamma_{p}\left(\alpha+\frac{1+p}{2}\right)|A|^{\alpha+\frac{1+p}{2}}\left\{\log |A|+\gamma_{p}\left(\alpha+\frac{1+p}{2}\right)\right\}\label{eq:intderivative}
\end{equation}
and the integral \eqref{eq:firstlogpart} can be computed by utilizing this result directly. To compute the second nontrivial log term \eqref{eq:2ndlogpart}
\begin{equation}
\int_{S\succ 0}j\left(T_{i},S,\left\{M,\Sigma,\Omega,n\right\}\right)\textrm{tr}\left[S(\Sigma_{t}+(T_{i}-M)\Omega^{-1}(T_{i}-M)^{\textrm{t}})\right]\,dS,
\end{equation}
we note that due to the linearity of the trace the result is equivalent to solving a matrix valued integral
\begin{equation}
\textrm{tr}\left[\left(\int_{S\succ 0}j\left(T_{i},S,\left\{M,\Sigma,\Omega,n\right\}\right)S\,dS\right)(\Sigma_{t}+(T_{i}-M)\Omega^{-1}(T_{i}-M)^{\textrm{t}})\right]\,.
\end{equation}
The enclosed integral also follows from Eq.~\eqref{eq:intmultgammadef}, after taking the matrix derivative with respect to $A$. Using, for symmetric $S$
\begin{equation}
\frac{d}{dA}\det A=(\det A)A^{-1}\label{eq:detderivative}
\end{equation}
and
\begin{equation}
\frac{d}{dA}e^{\textrm{tr}SA}=(e^{\textrm{trS}A})S
\end{equation}
we have 
\begin{equation}
\int_{S\succ 0}e^{-\textrm{tr}\left[SA\right]}|S|^{x}S\,dS=-\frac{1}{2}(1+p+2x)\Gamma_{p}\left(\frac{1}{2}(1+p+2x)\right)(\det A)^{\frac{1}{2}(p+2x-1)}A^{-1}
\end{equation}
from where again, we can obtain the result by direct substitution.

Combining results of the two integrals together with adequate rearrangements we obtain the integrated form of Eq.~\eqref{eq:expdef}
\begin{equation}
\begin{split}E=\sum_{i}^{N}n_{t}\psi_{p}\left(\frac{1}{2}(m+n+p-1)\right)-2\log\Gamma_{p}\left(\frac{1}{2}(n_{t}+p-1)\right)+p\log|\Omega_{t}|^{-1}\\
-n_{t}\log|(T_{i}-M)\Omega^{-1}(T_{i}-M)^{\textrm{t}}+\Sigma|+(n_{t}+p-1)\log |\Sigma_{t}|\\
-(m+n+p-1)\textrm{tr}\left(\left((T_{i}-M)\Omega^{-1}(T_{i}-M)^{\textrm{t}}+\Sigma\right)^{-1}\left((T_{i}-M_{t})\Omega_{t}^{-1}(T_{i}-M_{t})^{\textrm{t}}+\Sigma_{t}\right)\right)
\end{split}
\label{eq:fsimp}
\end{equation}
The maximization formulae is obtained setting the derivatives of this equation to zero.

>>>>

We can split the integral in Eq.~\eqref{eq:expdef} by the terms of $\log f(T_{j},S;\theta)$. Eliminating a global factor $\frac{1}{2}$ and terms which do not depend on the parameters in $\theta$ we have
\begin{equation}
\begin{split}\log f(T_{j},S;\theta)\sim -\nu\log|S|-\textrm{tr}\left[S^{-1}(\Sigma+(T_j-M)\Omega^{-1}(T_j-M)^{\textrm{t}})\right]\\
-\nu p\log2-p\log |\Omega|+(\nu+p-1)\log |\Sigma|-2\log\Gamma_{p}\left(\frac{\nu+p-1}{2}\right).
\end{split}\label{eq:logf}
\end{equation}
From Eq.~\eqref{eq:logf}, we can see the only integrals which need to be computed are
\begin{equation}
\int_{\mathbf{S}_+^p}f(T_j,S;\theta_k)\log |S|\,dS\label{eq:firstlogpart}
\end{equation}
and
\begin{equation}
\int_{\mathbf{S}_+^p}f(T_j,S;\theta_k) S^{-1}\,dS.\label{eq:2ndlogpart}
\end{equation}
Note that taking the derivative with respect to $\alpha$ in Eq.~\eqref{eq:intmultgammadef} we have
\begin{equation}
\int_{\mathbf{S}_+^p}e^{-\textrm{tr}\left[SA\right]}|S|^{\alpha}\log |S|\,dS=\Gamma_{p}\left(\alpha+\frac{1+p}{2}\right)|A|^{\alpha+\frac{1+p}{2}}\left\{\log |A|+\gamma_{p}\left(\alpha+\frac{1+p}{2}\right)\right\}\label{eq:intderivative1}
\end{equation}
and hence the integral \eqref{eq:firstlogpart} follows this result directly. To compute the integral \eqref{eq:2ndlogpart} we take the matrix derivative of Eq.~\eqref{eq:intmultgammadef} with respect to $A$ and obtain
\begin{equation}
\int_{\mathbf{S}_+^p}e^{-\textrm{tr}\left[SA\right]}|S|^{\alpha}S\,dS=-\left(\alpha+\frac{1+p}{2}\right)\Gamma_{p}\left(\alpha+\frac{1+p}{2}\right)|A|^{\alpha+\frac{p-1}{2}}A^{-1},\label{eq:intderivative2}
\end{equation}
from where again, we can obtain the result by direct substitution.

Combining results of Eqs.~\eqref{eq:intderivative1} and \eqref{eq:intderivative2} with adequate rearrangements we obtain the closed form of $Q_j(\theta,\theta_k)$ while discarding the irrelevant constant terms
\begin{equation}
\begin{split}Q_j(\theta,\theta_k)\sim \nu\psi_{p}\left(\frac{m+\nu_k+p-1}{2}\right)-2\log\Gamma_{p}\left(\frac{\nu+p-1}{2}\right)+p\log|\Omega|^{-1}\\
-\nu\log|(T_j-M_k)\Omega_k^{-1}(T_j-M_k)^{\textrm{t}}+\Sigma_k|+(\nu+p-1)\log |\Sigma|\\
-(m+\nu_k+p-1)\textrm{tr}\left[\left((T_j-M_k)\Omega_k^{-1}(T_j-M_k)^{\textrm{t}}+\Sigma_k\right)^{-1}\right.\\
\left.\left((T_j-M)\Omega^{-1}(T_j-M)^{\textrm{t}}+\Sigma\right)\right].
\end{split}
\label{eq:fsimp}
\end{equation}
The maximization formulae is obtained setting the derivatives of this equation to zero.
